---
title: "Summary of home range"
output: html_notebook
---
```{r, message=FALSE, warning=FALSE, echo=FALSE}
library('tidyverse')
library('lubridate')
library('move')
library('adehabitatHR')
library('knitr')
library('kableExtra')
library('sf')
library('ggplot2')
```

```{r, warning=FALSE, echo=FALSE}
# Read in the data.
df <- read.csv('../data/processed/telem_all.csv',
               header=TRUE, stringsAsFactors=FALSE) %>%
  drop_na('lat')

# Do the datetime thing.
df$datetime <- ymd_hms(df$datetime, tz='America/Vancouver')

# Create intervals.
winter2 <- interval(ymd(20190101), ymd(20190414))
incubation <- interval(ymd(20190415), ymd(20190510))
nestling <- interval(ymd(20190511), ymd(20190710))
fledgling <- interval(ymd(20190711), ymd(20190901))
winter1 <- interval(ymd(20190902), ymd(20191231))

# Assign locations to period.
df <- df %>%
  mutate(yrls=ymd(paste(2019, month(datetime), day(datetime))),
         period=case_when(
           yrls %within% incubation ~ 'incubation',
           yrls %within% nestling ~ 'nestling',
           yrls %within% fledgling ~ 'fledgling',
           yrls %within% winter1 ~ 'winter',
           yrls %within% winter2 ~ 'winter'
         )
  )

# Do the spatial thing.
sf.df <- st_as_sf(df, coords=c('lon', 'lat')) %>%
  st_set_crs('+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs') %>%
  st_transform("+proj=utm +zone=10 +datum=WGS84 +units=m +no_defs")

# 50% and 95% MCP for each tag and all points.
mcp.all <- sf.df %>%
  dplyr::select(geometry, id) %>%
  as_Spatial() %>%
  mcp.area(percent=c(50, 95),
           unin='m', unout='ha', plotit=FALSE)%>%
  rownames_to_column(var='name') %>%
  mutate(percent=as.numeric(name)) %>%
  dplyr::select(-name)

# Add new columns
mcp.all <- mcp.all %>%
  add_column(method='MCP', period='total')

# Fix formatting so it can actually be graphed.
mcp.all <- pivot_longer(mcp.all,
  cols=c('HAR04', 'HAR05', 'HAR07', 'HAR08', 'HAR09', 'HAR10'), names_to='id')

# 50% and 95% MCP for each tag and nestling points.
mcp.nest <- sf.df %>%
  filter(period == 'nestling') %>%
  dplyr::select(geometry, id) %>%
  as_Spatial() %>%
  mcp.area(percent=c(50, 95),
           unin='m', unout='ha', plotit=FALSE)%>%
  rownames_to_column(var='name') %>%
  mutate(percent=as.numeric(name)) %>%
  dplyr::select(-name)

mcp.nest <- mcp.nest %>%
  add_column(method='MCP', period='nestling') %>%
  pivot_longer(cols=c('HAR04', 'HAR05', 'HAR07', 'HAR08', 'HAR09', 'HAR10'),
               names_to='id')

# 50% and 95% MCP for each tag and fledgling points.
mcp.fledge <- sf.df %>%
  filter(period == 'fledgling') %>%
  dplyr::select(geometry, id) %>%
  as_Spatial() %>%
  mcp.area(percent=c(50, 95),
           unin='m', unout='ha', plotit=FALSE)%>%
  rownames_to_column(var='name') %>%
  mutate(percent=as.numeric(name)) %>%
  dplyr::select(-name)

mcp.fledge <- mcp.fledge %>%
  add_column(method='MCP', period='fledgling') %>%
  pivot_longer(cols=c('HAR05', 'HAR07'),
               names_to='id')

# 50% and 95% MCP for each tag and winter points.
mcp.wint <- sf.df %>%
  filter(period == 'winter') %>%
  dplyr::select(geometry, id) %>%
  as_Spatial() %>%
  mcp.area(percent=c(50, 95),
           unin='m', unout='ha', plotit=FALSE)%>%
  rownames_to_column(var='name') %>%
  mutate(percent=as.numeric(name)) %>%
  dplyr::select(-name)

mcp.wint <- mcp.wint %>%
  add_column(method='MCP', period='winter') %>%
  pivot_longer(cols=c('HAR05', 'HAR07'),
               names_to='id')

# Now make a kud.
kud.all <- sf.df %>%
  dplyr::select(id, geometry) %>%
  as_Spatial() %>%
  kernelUD(extent=0.25, grid=100) %>%
  kernel.area(percent=95, unin='m', unout='ha') %>%
  as.data.frame %>%
  rownames_to_column(var='name') %>%
  mutate(percent=as.numeric(name)) %>%
  dplyr::select(-name) %>%
  add_column(method='KDE', period='total') %>%
  pivot_longer(cols=c('HAR04', 'HAR05', 'HAR07', 'HAR08', 'HAR09', 'HAR10'),
               names_to='id')

# And for 50%.
k.a.50 <- sf.df %>%
  dplyr::select(id, geometry) %>%
  as_Spatial() %>%
  kernelUD(extent=0.25, grid=100) %>%
  kernel.area(percent=50, unin='m', unout='ha') %>%
  as.data.frame %>%
  rownames_to_column(var='name') %>%
  mutate(percent=as.numeric(name)) %>%
  dplyr::select(-name) %>%
  add_column(method='KDE', period='total') %>%
  pivot_longer(cols=c('HAR04', 'HAR05', 'HAR07', 'HAR08', 'HAR09', 'HAR10'),
               names_to='id')

# For nestling 95%.
k.n.95 <- sf.df %>%
  filter(period == 'nestling') %>%
  dplyr::select(geometry, id) %>%
  as_Spatial() %>%
  kernelUD(extent=0.3, grid=100) %>%
  kernel.area(percent=95, unin='m', unout='ha') %>%
  as.data.frame %>%
  rownames_to_column(var='name') %>%
  mutate(percent=as.numeric(name)) %>%
  dplyr::select(-name) %>%
  add_column(method='KDE', period='nestling') %>%
  pivot_longer(cols=c('HAR04', 'HAR05', 'HAR07', 'HAR08', 'HAR09', 'HAR10'),
               names_to='id')

# For nestling 50%.
k.n.50 <- sf.df %>%
  filter(period == 'nestling') %>%
  dplyr::select(geometry, id) %>%
  as_Spatial() %>%
  kernelUD(extent=0.3, grid=100) %>%
  kernel.area(percent=50, unin='m', unout='ha') %>%
  as.data.frame %>%
  rownames_to_column(var='name') %>%
  mutate(percent=as.numeric(name)) %>%
  dplyr::select(-name) %>%
  add_column(method='KDE', period='nestling') %>%
  pivot_longer(cols=c('HAR04', 'HAR05', 'HAR07', 'HAR08', 'HAR09', 'HAR10'),
               names_to='id')

# For fledgling 95%
k.f.95 <- sf.df %>%
  filter(period == 'fledgling') %>%
  dplyr::select(geometry, id) %>%
  as_Spatial() %>%
  kernelUD(extent=0.3, grid=100) %>%
  kernel.area(percent=95, unin='m', unout='ha') %>%
  as.data.frame %>%
  rownames_to_column(var='name') %>%
  mutate(percent=as.numeric(name)) %>%
  dplyr::select(-name) %>%
  add_column(method='KDE', period='fledgling') %>%
  pivot_longer(cols=c('HAR05', 'HAR07'),
               names_to='id')

k.f.50 <- sf.df %>%
  filter(period == 'fledgling') %>%
  dplyr::select(geometry, id) %>%
  as_Spatial() %>%
  kernelUD(extent=0.3, grid=100) %>%
  kernel.area(percent=50, unin='m', unout='ha') %>%
  as.data.frame %>%
  rownames_to_column(var='name') %>%
  mutate(percent=as.numeric(name)) %>%
  dplyr::select(-name) %>%
  add_column(method='KDE', period='fledgling') %>%
  pivot_longer(cols=c('HAR05', 'HAR07'),
               names_to='id')

# Merge everything.
hr <- bind_rows(mcp.all, mcp.nest, mcp.fledge, mcp.wint, kud.all, k.a.50, k.n.95, k.n.50, k.f.95, k.f.50)

hr <- hr %>%
  pivot_wider(names_from=period, values_from=value) %>%
  dplyr::select(id, method, percent, total, nestling, fledgling, winter) %>%
  arrange(id, method, percent)

# Create some summary stuff.
df.sum <- df %>%
  dplyr::select(id, period) %>%
  group_by(id, period) %>%
  summarize(n=n())

df.sum <- df.sum %>%
  pivot_wider(names_from=period, values_from=n, values_fill=list(n=0)) %>%
  mutate(total=incubation+nestling+fledgling+winter, percent=NA, 
         method='Number of points') %>%
  dplyr::select(id, method, percent, total, incubation, nestling, fledgling, winter)

# Add to main table.
hr2 <- bind_rows(hr, df.sum)

hr2 <- hr2 %>%
  arrange(id, method, percent) %>%
  dplyr::select(id, method, percent, total, incubation, nestling, fledgling, winter)
```
In June of 2018 and 2019, I attached [^1] GPS geologgers to 6 northern goshawks in south coastal B.C. Based on dates from the literature, I divided the annual breeding cycle into the four periods (see table).
```{r warning=FALSE, echo=FALSE}
Period <- c('Incubation', 'Nestling', 'Fledgling', 'Winter')
Start <- c(ymd(20190415), ymd(20190511), ymd(20190711), ymd(20190902))
End <- c(ymd(20190510), ymd(20190710), ymd(20190901), ymd(20190414))
pers <- data.frame(Period, Start, End)

kable(pers) %>%
  kable_styling(bootstrap_options='condensed', full_width=F, position='left')
```
Using `adehabitatHR`, I calculated 50% and 95% home ranges for all points and for each period of the breeding season, where sufficient location points were available, using both MCP and KDE methods. For the UD estimates, I used the default bandwith, as the least-squares method is not very good for "clumpy" points.
```{r warning=FALSE, echo=FALSE}
options(knitr.kable.NA='')
kable(hr2, digits=2, align='c',
      col.names=c('ID', 'Method', '%', 'Total (ha)', 'Incubation', 'Nestling',
                 'Fledgling', 'Winter')) %>%
  kable_styling(full_width = T) %>%
  column_spec(1, bold = T) %>%
  collapse_rows(columns=1:2, valign='top') %>%
  row_spec(c(5, 10, 15, 20, 25, 30), bold=T, background='lightgray') %>%
  add_header_above(c(" "=4, "Period"=4))
```
This table is missing KDE-method home ranges for those birds with winter points (HAR05 and HAR07). Note the sample size--some estimates are more reliable than others due to rather small numbers of points. Because all birds have some points for the nestling season, I made a summary chart of home range estimates based on all available points and on nestling season points only.
```{r  warning=FALSE, echo=FALSE}
ggplot(hr, aes(x=factor(id), y=total)) +
  geom_jitter(aes(shape=factor(method), color=factor(percent)), size=3, alpha=0.7,
              position=position_jitter(width=0.2)) +
  theme_classic() +
  labs(x='ID', y='Home range size (ha)', title='Total home range',
  caption='Home range size calculated with all location points available in 2019') +
  guides(color=guide_legend('Percent'), shape=guide_legend('Method'))

ggplot(hr, aes(x=factor(id), y=nestling)) +
  geom_jitter(aes(shape=factor(method), color=factor(percent)), size=3, alpha=0.7,
              position=position_jitter(width=0.2)) +
  theme_classic() +
  labs(x='ID', y='Home range size (ha)', title='Nestling season home range',
       caption='Home range size calculated with nestling season (11 May-10 July) points available in 2019') +
  guides(color=guide_legend('Percent'), shape=guide_legend('Method'))
```
These two graphs are very similar, though the nestling season graph drops the relative size of HAR07's home range. The KDE and MCP estimates are also generally very similar, which is perhaps not surprising in a central-place forager that uses area pretty evenly around its nest.
[^1]: And by "I" I mean my awesome colleagues and helpers did most of the hard work because they are amazing!
