---
title: "Spatial storytime"
output: html_notebook
---

# Summary of available telemetry data

```{r options, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE)
```

```{r message=FALSE}
# Import conflict settings.
source('../src/conflicted.R')

#Load some libraries.
library(tidyverse)
library(sf)
library(lubridate)
library(adehabitatHR)
library(ggplot2)
library(extrafont)

# Define some colors.
m <- '#d36135' # Flame
f <- '#689689' # Polished pine

# Load telemetry data.
tl <- read_csv('../data/processed/telemetry_2018-2019.csv')

# RLK appears to have been missexed. I'll change that here.
tl <- tl %>% mutate(sex = case_when(
  id == 'HAR04' ~ 'm',
  TRUE ~ sex
  ))
```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
