---
title: "Telemetry refresher"
output: html_notebook
---

```{r options, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, message=FALSE, warning=FALSE)
```

Wow so I haven't looked at any telemetry anything since April. And now I am supposed to talk intelligently about it during my committee meeting.

```{r message=FALSE}
# Import conflict settings.
source('../src/conflicted.R')

#Load some libraries.
library(tidyverse)
library(sf)
library(lubridate)
library(adehabitatHR)
library(ggplot2)
library(extrafont)

# Define some colors.
m <- '#d36135' # Flame
f <- '#689689' # Polished pine

# Load telemetry data.
tl <- read_csv('../data/processed/telemetry_2018-2019.csv')

# RLK appears to have been missexed. I'll change that here.
tl <- tl %>% mutate(sex = case_when(
  id == 'HAR04' ~ 'm',
  TRUE ~ sex
  ))
```

Start with some summary information. Note that there was another female tagged in 2019, UTZ, for which no data has been retrieved.

```{r}
# How many tags?
tl %>% distinct(site, year(date), id, sex)

# How much data per tag?
tl %>% group_by(id) %>%
  mutate(min=min(date), max=max(date), n.points=n(), period=max-min) %>% 
  distinct(id, site, n.points, min, max, period) 
```

These points span some pretty different periods. To try to standardize them, I can potentially chose just points within certain stages of the breeding season.

```{r warning=FALSE}
# Define breeding season.
breeding.2018 <- interval(ymd(20180511), ymd(20180901))
breeding.2019 <- interval(ymd(20190511), ymd(20190901))


# Select only points that fall within the breeding season.
tl.breeding <- tl %>% 
  filter(date %within% c(breeding.2018, breeding.2019))

# Summarize.
tl.breeding %>% group_by(id) %>% 
  mutate(min=min(date), max=max(date), n.points=n(), period=max-min) %>% 
  distinct(id, site, n.points, min, max, period) 
```

These intervals combine what I previously called, separately, nestling and fledgling. Since I previously calculated home ranges using those separate categories, here I'll calculate total breeding-season home range.

```{r warning=FALSE}
# Transform telemetry data to a spatial format.
tl.breeding.sf <- tl.breeding %>% st_as_sf(coords=c('lon', 'lat')) %>%
  st_set_crs('+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs') %>%
  st_transform("+proj=utm +zone=10 +datum=WGS84 +units=m +no_defs")

# Make 95% & 50% MCPs
mcp <- tl.breeding.sf %>%
  select(id, geometry) %>%
  as_Spatial() %>%
  mcp.area(percent=c(50, 95), unin='m', unout='ha', plotit=FALSE) %>% 
  rownames_to_column(var='percent') %>% 
  pivot_longer(-percent, names_to='id', values_to='area') %>% 
  mutate(method='mcp')

# Make 95% and 50% KDEs.
kde <- tl.breeding.sf %>%
  select(id, geometry) %>%
  as_Spatial() %>%
  kernelUD() %>% 
  kernel.area(percent=c(50, 95), unin='m', unout='ha') %>% 
  rownames_to_column(var='percent') %>% 
  pivot_longer(-percent, names_to='id', values_to='area') %>% 
  mutate(method='kde')

# And in the darkness bind them.
homerange <- bind_rows(mcp, kde)

# Add sex info.
homerange <- tl %>% distinct(id, sex) %>% right_join(homerange, by=c('id'))
```

As a check of the methods, let's see how the KDEs and MCPs compare at different scales.

```{r}
homerange %>% mutate(group=paste(percent, method, sep='')) %>% 
  ggplot(aes(x=group, y=area)) +
  geom_boxplot() +
  theme_classic()
```

Visually, they look pretty similar. I think I'll be going with the MCPs even though they're stupid because a) they're more widely used and so more comparable and b) the KDE has some issues with the smaller homeranges.

More interestingly, let's look at the difference between males and females.

```{r}
homerange %>% filter(method == 'mcp') %>%
  ggplot(aes(x=as.factor(percent), y=area, fill=sex)) +
  geom_boxplot() +
  labs(title='Home range and core-use area by sex', x='% MCP', y='Area (ha)') +
  scale_fill_manual(values=c(f, m)) +
  theme_classic() +
  theme(text=element_text(family="Lato Semibold"))
```

Oh, yeah, huge difference. In table form:

```{r}
homerange %>% filter(method == 'mcp') %>% 
  group_by(sex, percent) %>% 
  summarize(n=n(), mean.area=mean(area))
```

By the way, is there any difference in home range size between coastal and transition zone birds?