---
title: "Image processing setup"
output: html_notebook
---

I have a ton of photos from nest cameras, and I need to figure out how to process them efficiently.
```{r echo=TRUE, message=FALSE, warning=FALSE}
# Load some libraries.
library('tidyverse')
library('exiftoolr')
library('lubridate')
library('purrr')

# This installs exiftool, which lets me view EXIF data.
# install_exiftool()

# Make a list of all the files in the directory.
TMC.files <- list.files('D:/RECONYX/TwentyMileCreek/100RECNX_TwentyMile', full.names=TRUE)

# Extract EXIF data from the files and make a dataframe.
TMC.photos <- exif_read(TMC.files,
                    tags=c('filename', 'CreateDate', 'SerialNumber'))

# Make it pretty.
photos <- TMC.photos %>%
  dplyr::select(filename=FileName, datetime=CreateDate,
                serial=SerialNumber) %>%
  mutate(site='TMC')

# Do the datetime thing.
photos$datetime <- ymd_hms(photos$datetime, tz='America/Vancouver')

# When was it recording?
photos %>% summarize(first=min(datetime), last=max(datetime),
          length=difftime(last, first))
```
That's pretty awesome because we were worried the camera memory would fill or the batteries would run out, but it looks like this one recorded the whole time.
```{r echo=TRUE, message=FALSE, warning=FALSE}
# How many photos per day?
photos %>%
  mutate(datetime=floor_date(datetime, unit='days')) %>%
  group_by(datetime) %>%
  summarize(n())
```
Well shoot, that's a lot of photos. Let's try this whole process with my "sample" photo set so I don't screw stuff up.
```{r echo=TRUE, message=FALSE, warning=FALSE}
# Make a list of all the files in the directory.
sample.files <- list.files('../data/raw/TMC', full.names=TRUE)

# Extract EXIF data from the files and make a dataframe.
sample.exif <- exif_read(sample.files,
                    tags=c('filename', 'CreateDate', 'SerialNumber'))

# Make it pretty.
sample.photos <- sample.exif %>%
  dplyr::select(filename=FileName, datetime=CreateDate,
                serial=SerialNumber) %>%
  mutate(site='TMC')

# Export as csv.
write.csv(sample.photos, file='../data/interim/sample_photos.csv',
          row.names=FALSE)
```
I took the csv, loaded it in sheets, and filled it out. Then I re-loaded it into R to see how it works.
```{r echo=TRUE, message=FALSE, warning=FALSE}
# Load in the filled csv.
# Select only those images that are "interesting."
df <- read.csv('../data/processed/sample_photos_cooked.csv', header=TRUE,
               stringsAsFactors=FALSE) %>%
  filter(interest != '')

# Get the filenames of the interesting photos.
interest <- sample.exif %>%
  filter(FileName %in% df$filename) %>%
  dplyr::select(SourceFile)

# Transfer the interesting photos to a new folder.
file.copy(from=interest$SourceFile,
          to='../data/raw/TMC/interest', overwrite=TRUE)
```
That seems to work! To go forward for real, then, I need to split my real csv into reasonable chunks for filling.
```{r echo=TRUE, message=FALSE, warning=FALSE}
# Add columns to dataframe for data entry.
photos <- photos %>%
  add_column(interest='', live.chicks='', class='', family='',
             genus='', species='', common='', comments='')

# Make a folder for all this.
# Happens just once.
# dir.create('../data/interim/n_cam_csv')

# Save the big csv, just in case.
write.csv(photos, file='../data/interim/n_cam_csv/TMC_photos.csv',
          row.names=FALSE)

# Now split the data frame by week.
cut <- split(df, week(df$date))

# And save each week at its own csv.
walk(seq_along(cut), function(x){
  write.csv(cut[[x]],
           file=paste('../data/interim/n_cam_csv/', 'TMC_photos', '_',
                      as.character(x), '.csv', sep=''),
           row.names=FALSE)
})

```