---
title: "Unrecovered data"
output: html_notebook
---

```{r options, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, message=FALSE, warning=FALSE)
```

Trying to see how much data we do have compared to how much data we should have.

```{r message=FALSE, warning=FALSE}
# Import conflict settings.
source('../src/conflicted.R')

#Load some libraries.
library(tidyverse)
library(sf)
library(ggplot2)
library(extrafont)

# Load telemetry data.
source('../src/clean_telemetry.R')
```

I won't bother doing this with number of points, but I will look at days, and maybe break it down in breeding season days and winter days.

```{r}
# Transform date into a date.
telemetry.sites <- telemetry.sites %>% mutate(date=ymd(date))

# Twist wide so each id gets its own row.
captures <- select(telemetry.sites, m_tag, f_tag, year, date, site, nest) %>%
  pivot_longer(!c(year, date, site, nest), names_to='sex', values_to='id') %>%
  drop_na(id) %>%
  mutate(sex=case_when(
    sex == 'm_tag' ~ 'm',
    sex == 'f_tag' ~ 'f'
  )) %>% 
  distinct(nest, id, .keep_all=TRUE)

# Get today's date.
now <- today()

# Calculate difference between tagging date and today's date.
captures <- captures %>% mutate(e.time=case_when(
  year == 2018 ~ ymd('2018-12-31') - date,
  TRUE ~ now - date))
```

Note that HAR07 died sometime in the winter of 2018, so it isn't fair to count all of the elapsed time since he was tagged--it''s cut off at the end of 2018.

Now calculate how many days of data there actually are. Again, I'll trim HAR07's data to the end of 2018.

```{r}
# Trim HAR07 points.
df <- df %>% filter(id != 'HAR07' | datetime <= ymd('2018-12-31'))

# Calculate days of data.
avail.data <- df %>% group_by(id) %>% 
  mutate(min=min(date), max=max(date), n.points=n(), period=max-min) %>% 
  distinct(id, site, n.points, min, max, period) 

# Then join them together.
t.data <- select(avail.data, id, site, period) %>% right_join(captures, by=c('site', 'id')) %>% 
  mutate(period=as.numeric(period), e.time=as.numeric(e.time)) %>% 
  replace_na(list(period=0))

# Calculate percentage of days retrieved.
t.data %>% mutate(p.time=period/e.time*100)
```

Ok, so that's per bird. How about overall?

```{r}
t.data %>% ungroup() %>% summarize(period=sum(period), e.time=sum(e.time)) %>% 
  mutate(p.time=period/e.time*100)
```


Ooof. So we have only retrieved about 10% of the days of data we should have. Most of that's during the winter, though, yes? Well, at this point maybe not...

```{r}
# Define breeding season.
breeding.2018 <- interval(ymd(20180420), ymd(20180915))
breeding.2019 <- interval(ymd(20190420), ymd(20190915))
breeding.2020 <- interval(ymd(20200420), ymd(20200915))

# Select only points that fall within the breeding season.
df.breeding <- df %>% 
  filter(date %within% c(breeding.2018, breeding.2019, breeding.2020))

# Calculate days of data.
avail.data.br <- df.breeding %>% group_by(id) %>% 
  mutate(min=min(date), max=max(date), n.points=n(), period=max-min) %>% 
  distinct(id, site, n.points, min, max, period) 

# Then join them together.
t.data.br <- select(avail.data.br, id, site, period) %>% right_join(captures, by=c('site', 'id')) %>% 
  mutate(period=as.numeric(period), e.time=as.numeric(e.time)) %>% 
  replace_na(list(period=0))

# Calculate percentage of days retrieved.
t.data.br %>% mutate(p.time=period/e.time*100)
```

And overall?

```{r}
t.data.br %>% ungroup() %>% summarize(period=sum(period), e.time=sum(e.time)) %>% 
  mutate(p.time=period/e.time*100)
```

Oh, wow, worse. How does that math even work??

The obvious question is *where* are all these data going? The most depressing option is tag mortality. 

* Blakely etal 2019 had 3/20 birds disappear over 3 years, using the same GPS logger tech and backpack mounts we used.
* Beier & Drennan 1999 lost 0/20 birds over 2 years, using VHF and backpacks. They report no mortality but do not explicitly state there were no losses.
* Boal etal 2003 lost 2/33 birds over 2 years wearing VHF and backpacks to predation. An additional bird lost its tag. (Also note that 3 females could not be included bc they had too few points.)
* McCLaren 2005 13/63 birds in first year after tagging (though study lasted 5 years) with VHF and backpacks. Did not get enough locations to generate a single homerange estimate.
* Squires & Ruggiero 1995 lost 3/4 birds tagged with VHF and backpacks. This was specifically winter study so low survival expected.

Actual studies of survival are variable. Steenhof etal 2006 found survival nearly halved for tagged prairie falcons, but Sergio etal 2015 found no significant effect in black kites.

Survival on Kaibab was has (Reynolds & Joy 1998), 0.688 for males and 0.866 for females. They also found territory turnover--a territory holder being replaced in a subsequent year--to be 25% for males and 16% for females.

So what actually happened to our birds?

```{r}
fate <- c('breeding 2020, no data', 
          'breeding 2020', 
          'breeding 2020', 
          'present at site, no 2020 data', 
          'present at site, no 2020 data',
          'breeding 2020, no data',
          'deceased',
          'breeding 2020, no data',
          'potentially resident at unknown nest',
          'present at new site, no 2020 data',
          'present at site, no 2020 data',
          'unknown, not breeding at site',
          'breeding 2020')

t.data %>% arrange(id) %>% bind_cols(fate=fate) %>% select(!date) %>% arrange(year)
```

Not actually as bad as it sounds.


